Biomarker screening

### 2022-04-22 code/r/02_C_numeric_vs_hpo_for_blood_disorder.Rmd
update

### retrieve data
```{r message=FALSE}
library(tidyverse)
library(glue)
library(lubridate)
library(readxl)
require(DBI)
require(RPostgres)
require(egg)
require(gridExtra)
source('.db_credential.R')
source('functions.R')
dsca <- DBI::dbConnect(RPostgres::Postgres(), host='localhost', user=user, dbname = dbname, port = 2345, password=password)

# cache directory to read from or write to
CACHE_DIR <- './cache'
if (!dir.exists(CACHE_DIR)){
  dir.create(CACHE_DIR)
}

FORCE_RERUN <- TRUE
```

>>>>

### retrieve data
```{r message=FALSE}
library(tidyverse)
library(glue)
library(lubridate)
library(readxl)
require(DBI)
require(RPostgres)
require(egg)
require(gridExtra)
source('functions.R')
dsca <- DBI::dbConnect(RPostgres::Postgres(), 
                       host=Sys.getenv('DSCA_PHI_HOST'), 
                       dbname = Sys.getenv('DSCA_PHI_DBNAME'), 
                       port = Sys.getenv('DSCA_PHI_PORT'), 
                       user=Sys.getenv('DSCA_PHI_USERNAME'),
                       password= Sys.getenv('DSCA_PHI_PASSWORD'))

# cache directory to read from or write to
CACHE_DIR = '../../cache'

if (!dir.exists(CACHE_DIR)){
  dir.create(CACHE_DIR)
}

FORCE_RERUN <- FALSE
```

### 2022-04-22 code/r/02_D_numeric_vs_hpo_for_CRC.Rmd
update

### retrieve data
```{r message=FALSE}
library(tidyverse)
library(glue)
library(lubridate)
library(readxl)
require(DBI)
require(RPostgres)
require(egg)
require(gridExtra)
source('.db_credential.R')
source('functions.R')
dsca <- DBI::dbConnect(RPostgres::Postgres(), host='localhost', user=user, dbname = dbname, port = 2345, password=password)

# cache directory to read from or write to
CACHE_DIR <- './cache'
if (!dir.exists(CACHE_DIR)){
  dir.create(CACHE_DIR)
}
FORCE_RERUN <- TRUE
```

>>>>

### retrieve data
```{r message=FALSE}
library(tidyverse)
library(glue)
library(lubridate)
library(readxl)
require(DBI)
require(RPostgres)
require(egg)
require(gridExtra)
source('functions.R')
dsca <- DBI::dbConnect(RPostgres::Postgres(), 
                       host=Sys.getenv('DSCA_PHI_HOST'), 
                       dbname = Sys.getenv('DSCA_PHI_DBNAME'), 
                       port = Sys.getenv('DSCA_PHI_PORT'), 
                       user=Sys.getenv('DSCA_PHI_USERNAME'),
                       password= Sys.getenv('DSCA_PHI_PASSWORD'))

# cache directory to read from or write to
CACHE_DIR = '../../cache'

if (!dir.exists(CACHE_DIR)){
  dir.create(CACHE_DIR)
}
FORCE_RERUN <- TRUE
```

### 2022-04-22 code/r/02_E_numeric_vs_hpo_combined.Rmd
update

```{r}
library(tidyverse)
library(glue)
```

Load previous data analysis results
```{r}
liver_analysis <- read.csv('cache/liver_analysis.csv')
crc_analysis <- read.csv('cache/crc_case_matched_controls_analysis.csv')
acute_kidney_failure_analysis <- read.csv('cache/acute_kidney_failure_analysis.csv')
anemia_analysis <- read.csv('cache/anemia_analysis.csv')
```

@@

ggsave(filename = 'cache/plot_AUC_differences_numeric_vs_hpo.pdf', width = 6, height = 6)

>>>>

```{r}
library(tidyverse)
library(glue)

# cache directory to read from or write to
CACHE_DIR = '../../cache'
```

Load previous data analysis results
```{r}
liver_analysis <- read.csv(glue('{CACHE_DIR}/liver_analysis.csv'))
crc_analysis <- read.csv(glue('{CACHE_DIR}/crc_case_matched_controls_analysis.csv'))
acute_kidney_failure_analysis <- read.csv(glue('{CACHE_DIR}/acute_kidney_failure_analysis.csv'))
anemia_analysis <- read.csv(glue('{CACHE_DIR}/anemia_analysis.csv'))
```

@@

ggsave(filename = glue('{CACHE_DIR}/plot_AUC_differences_numeric_vs_hpo.pdf'), width = 6, height = 6)

### 2022-04-22 code/r/03_biomarker.Rmd
update

```{r message=FALSE}
library(tidyverse)
library(glue)
library(readxl)
require(DBI)
require(RPostgres)
require(egg)
require(gridExtra)
require(lubridate)
require(furrr)
source('.db_credential.R')
dsca <- DBI::dbConnect(RPostgres::Postgres(), host='localhost', user=user, dbname = dbname, port = 2345, password=password)
```

@@

```{sql connection=dsca, output.var=cohort}
select * from cohort;
```

```{sql connection=dsca, output.var=tested_phenotypes}
with testable_phenotypes_all as (
	select lab.medical_record_number , lab.context_name , lab.lab_date , lab.lab_time , lab.loinc , testable.termid , testable.is_tested_directly 
	from hai_az_prod.scc_and_epic_lab_after_loinc2hpo_2020july lab
	join hai_az_prod.loinc_testable_phenotypes testable on lab.loinc = testable.loincid 
	join cohort using (medical_record_number)
  )
 select medical_record_number, lab_date, termid
 from testable_phenotypes_all;
```

```{sql connection=dsca, output.var=observed_phenotypes}
with abnormal_lab as (
  select lab.*
  from hai_az_prod.scc_and_epic_lab_after_loinc2hpo_2020july lab 
  join cohort using (medical_record_number)
  where hpotermid is not null and not isnegated),
abnormal_lab_after_infer as (
  SELECT l.medical_record_number, l.lab_date, date_part(year, lab_date) as lab_year, l.local_test_code, l.loinc, l.hpotermid, l.isnegated, pair.ancestor , pair.distance
  FROM abnormal_lab l 
  left join hai_az_prod.hpo_is_a_pairs pair on l.hpotermid = pair.current)
select medical_record_number , lab_date, ancestor as termid
from abnormal_lab_after_infer;
```

```{sql connection=dsca, output.var='cohort_demographics'}
select person.*
from hai_az_prod.person_msdw_2020july person
join cohort using (medical_record_number);
```

@@

Annotate the screen results by adding the HPO term label, and also create a few different views that rank the results differently. 
```{r}
hpo_term_list <- read.csv('/Users/xingminzhang/git/s4-pd-dmsdw/pd-dmsdw-loinc2hpo-app/src/main/resources/hp_term_list.csv', header = FALSE, col.names = c('termid', 'distance_to_root', 'label'), stringsAsFactors = FALSE)

@@

write.csv(screen_results_annotated, file = '../cache/screen_results_annotated.csv', row.names = FALSE)

@@

write.csv(screen_results_annotated_formatted, file = '../cache/screen_results_annotated.csv', row.names = FALSE)

@@

sjPlot::tab_df(eosinophilia %>% rename(`Lab Year Bracket` = lab_date_rel_bracket), 
                title = 'Table Association of Eosinophilia with Severe Asthma',
               file = '../cache/table_eosinophilia_with_severe_asthma.html',
               alternate.rows = TRUE)

@@

```{r}
sjPlot::tab_df(significant_by_tested_and_observed_deduped %>% select(-distance_to_root, -lab_date_rel_bracket), 
                title = 'Table Selected Phenotypes (within 1 year prior) Associated with Severe Asthma',
               file = '../cache/table_selected_biomarker_from_labs_within_1_year_prior_to_dx.html', 
               alternate.rows = TRUE)
sjPlot::tab_df(significant_by_tested_and_observed_deduped_three_years %>% 
                 select(-distance_to_root) %>% rename(`Lab Year Bracket` = lab_date_rel_bracket), 
                title = 'Table Selected Phenotypes (within past 3 years) Associated with Severe Asthma',
               file = '../cache/table_selected_biomarker_from_labs_within_3_years_prior_to_dx.html',
               alternate.rows = TRUE)
```

@@

  for (i in 1:length(cache_list)){
    filename <- names(cache_list)[[i]]
    write.table(cache_list[[i]], file = glue('../cache/{filename}.csv'), sep = ',', row.names = FALSE, col.names = TRUE)
  }
}
```

>>>>

```{r message=FALSE}
library(tidyverse)
library(glue)
library(readxl)
require(DBI)
require(RPostgres)
require(egg)
require(gridExtra)
require(lubridate)
require(furrr)
require(dbplyr)

dsca <- DBI::dbConnect(RPostgres::Postgres(), 
                       host=Sys.getenv('DSCA_PHI_HOST'), 
                       dbname = Sys.getenv('DSCA_PHI_DBNAME'), 
                       port = Sys.getenv('DSCA_PHI_PORT'), 
                       user=Sys.getenv('DSCA_PHI_USERNAME'),
                       password= Sys.getenv('DSCA_PHI_PASSWORD'))

# cache directory to read from or write to
CACHE_DIR = '../../cache'

if (!dir.exists(CACHE_DIR)){
  dir.create(CACHE_DIR)
}
FORCE_RERUN <- TRUE
```

@@

Get data from the database. The following method are lazily executed.
```{r}
cohort <- tbl(dsca, 'cohort')
tested_phenotypes <- tbl(dsca, sql("
    with testable_phenotypes_all as (
    	select lab.medical_record_number , lab.context_name , lab.lab_date , lab.lab_time , lab.loinc , testable.termid , testable.is_tested_directly
    	from hai_az_prod.scc_and_epic_lab_after_loinc2hpo_2020july lab
    	join hai_az_prod.loinc_testable_phenotypes testable on lab.loinc = testable.loincid
    	join cohort using (medical_record_number)
      )
     select medical_record_number, lab_date, termid
     from testable_phenotypes_all"))
observed_phenotypes <- tbl(dsca, sql("
    with abnormal_lab as (
      select lab.*
      from hai_az_prod.scc_and_epic_lab_after_loinc2hpo_2020july lab
      join cohort using (medical_record_number)
      where hpotermid is not null and not isnegated),
    abnormal_lab_after_infer as (
      SELECT l.medical_record_number, l.lab_date, date_part(year, lab_date) as lab_year, l.local_test_code, l.loinc, l.hpotermid, l.isnegated, pair.ancestor , pair.distance
      FROM abnormal_lab l
      left join hai_az_prod.hpo_is_a_pairs pair on l.hpotermid = pair.current)
    select medical_record_number , lab_date, ancestor as termid
    from abnormal_lab_after_infer"))
cohort_demographics <- tbl(dsca, sql("
    select person.*
    from hai_az_prod.person_msdw_2020july person
    join cohort using (medical_record_number)"))

```

@@

Annotate the screen results by adding the HPO term label, and also create a few different views that rank the results differently. 
```{r}
hpo_term_list <- read.csv('../../in/hp_term_list.csv', header = FALSE, col.names = c('termid', 'distance_to_root', 'label'), stringsAsFactors = FALSE)

@@

write.csv(screen_results_annotated, file = glue('{CACHE_DIR}/screen_results_annotated.csv'), row.names = FALSE)

@@

write.csv(screen_results_annotated_formatted, file = glue('{CACHE_DIR}/screen_results_annotated.csv', row.names = FALSE)

@@

sjPlot::tab_df(eosinophilia %>% rename(`Lab Year Bracket` = lab_date_rel_bracket), 
                title = 'Table Association of Eosinophilia with Severe Asthma',
               file = glue('{CACHE_DIR}/table_eosinophilia_with_severe_asthma.html',
               alternate.rows = TRUE)

@@

```{r}
sjPlot::tab_df(significant_by_tested_and_observed_deduped %>% select(-distance_to_root, -lab_date_rel_bracket), 
                title = 'Table Selected Phenotypes (within 1 year prior) Associated with Severe Asthma',
               file = glue('{CACHE_DIR}/table_selected_biomarker_from_labs_within_1_year_prior_to_dx.html', 
               alternate.rows = TRUE)
sjPlot::tab_df(significant_by_tested_and_observed_deduped_three_years %>% 
                 select(-distance_to_root) %>% rename(`Lab Year Bracket` = lab_date_rel_bracket), 
                title = 'Table Selected Phenotypes (within past 3 years) Associated with Severe Asthma',
               file = glue('{CACHE_DIR}/table_selected_biomarker_from_labs_within_3_years_prior_to_dx.html',
               alternate.rows = TRUE)
```

@@

  for (i in 1:length(cache_list)){
    filename <- names(cache_list)[[i]]
    write.table(cache_list[[i]], file = glue('{CACHE_DIR}/{filename}.csv'), sep = ',', row.names = FALSE, col.names = TRUE)
  }
}
```
